Convolutional Neural Network to Detect Equivalent Questions in Online Forums \\ DL-IC 2018 Project

### 2018-07-04 paper/paper_final.tex
Dataset paper

%%%%%%%%% BODY TEXT
\section{Introduction}
Natural Language Processing (NLP) is just one of several fields emerging from Artificial Intelligence. In NLP we want the machine to process and analyze large amount of natural language data to accomplish very interesting tasks (such as Sentiment Analysis, Language Identification and many others). Nowadays, these tasks are plenty and many works on them has been already carried out \footnote{Here is a very interesting repository collecting several NLP tasks along with their main references: \\ \url{https://github.com/Kyubyong/nlp_tasks}}. In this paper, we focus on the task of predicting whether two questions from an online forum are semantically equivalent, i.e. they can be answered by the exact same answer. This can be a very useful tool for Question-Answering community sites as they typically want to keep their databases as less redundant as possible, so to make the searches of their users both faster and more effective. Furthermore, they want to keep people from wasting their time in answering already solved questions. \\
The main contribution of this work is \cite{bogdanova2015detecting}. In their work, Bogdanova et al. propose a simple Convolutional Neural Network (CNN) to detect semantically equivalent questions. The proposed network first transforms words into \emph{word embeddings} and then applies a convolutional layer which generates question-wide vectors representations. Finally, pairs of questions are scored using a similarity function. The most interesting experimental result is the evaluation of in-domain word embeddings versus the word embeddings generated using all of the English Wikipedia. In particular, they shown that domain-specific word embeddings achieve higher performances. \\
In general, the results obtained by our implementation are very close to the ones of the replicated paper. Surprisingly, as we will discuss in Section 4, the main differences in the results are about the optimal hyper-parameter values of the network.\\
The authors of the paper accurately described the architecture of the network, but many crucial details of the model are missing in their description. Hence, we have found quite stimulating to replicate this paper filling the holes with our own ideas. In particular, our efforts focused mainly on the following three aspects that were not explicitly treated in the paper: 
    \paragraph{Handling questions of different sizes.}
    Bogdanova et al. explicitly state that the different sizes of questions is one of the main challenges of their work. Nevertheless, they do not specify the way they solved this issue. We, therefore, tried to fix the length of each question to the \textbf{maximum} and to the \textbf{mean} length of all the questions. The first approach resulted to yield the greater accuracy.
    \paragraph{The type of convolution.}
    For the implementation of our network we used a 1-dimensional convolution. Indeed, text can be seen as images with the input maps having only one dimension (height, without width) corresponding to the question length and where the number of input maps is nothing but the size of the word embeddings. The type of convolution represents another missing detail of \cite{bogdanova2015detecting}.    
    \paragraph{The training of in-domain word embeddings.}
    The paper does not provide any information about the model used to train in-domain word embeddings. Initially, we tried to build our own \emph{word2vec} model \cite{mikolov2013distributed} to generate word-embeddings using as training data the same dataset used to train the main network \footnote{To do so we followed this good tutorial: \url{http://adventuresinmachinelearning.com/word2vec-keras-tutorial/}}. However, the training of this network turned out to be very slow and resulting embeddings perfomed poorly. \\
    We, then, decided to use Gensim, a very simple library which generates word-embeddings given the training data. The resulting embeddings performed very well, allowing the main network to achieve up to 90\% of accuracy on the validation set.

@@

%-------------------------------------------------------------------------
\subsection{Question of different length}
As starting point let's understand why unequal text length is a problem in NLP. The basic recipe for processing text in neural models is as follows:
\begin{itemize}\itemsep0.1pt
\item Convert text to tokens
\item Convert tokens to token IDs
\item Look up embedding vectors for IDs and construct a 2D tensor that are input to subsequent layers.
\end{itemize}•
Now, most neural network APIs (such as Keras) are “define-and-run” so we cannot change the dimensions of various matrices and tensors during runtime. All the techniques that can be used to tackle this issue force us to introduce new hyper-parameters or work-around in our model which are not fundamental for  the NLP task but are present to only handle unequal text lengths.\\
We decided to try two of these techniques, which we found out to be the best practices:
\begin{itemize}
	\item Fixing the length of all the questions to the \textbf{maximum} length. All the questions that are shorter are padded using the \emph{UNKNOWN} word (index $0$ of the vocabulary).
	\item Fixing the length of all the questions to the \textbf{mean} length. All the questions that are shorter are padded as in the case above; all the questions that are longer are chopped.
\end{itemize}
To keep the architecture of our network as simpler as possible we decided to apply this techniques as a preprocessing phase, without adding any further layer to the network architecture. Therefore, our network accepts as input only questions of the same length.\\
The experimental results show that the \emph{max} approach provides the best performances. This is probably due to the fact that the \emph{mean} approach is very likely to drop important semantic information from the sentence. It is worth mentioning that, in order for the \emph{max} approach to perform at its best, it can be a good choice dropping the outliers, i.e. the questions that are too long with respect to the other observations.
%-------------------------------------------------------------------------
\subsection{Type of convolution}
To implement the convolution described in Section 3.1.2., we decided to use a 1-dimensional convolutional layer. The reason behind this choice is that, as images are seen as three input maps (R,G,B) where each map is a 2-dimensional array (matrix) with height and width, similarly text can be seen as $d$ input maps where each map has only one dimension which is the length of the question. This interpretation of the 1-dimensional convolution is exemplified in Figure \ref{fig:conv}.\\
After applying vertically the sum operation to the output of this layer, the resulting vector is exactly the one described by Equation \ref{eq:conv}

@@

\paragraph{Datasets.}

>>>>

%%%%%%%%% BODY TEXT
%------------------------------------------------------------------------
\section{Introduction}
Natural Language Processing (NLP) is just one of several fields emerging from Artificial Intelligence. In NLP we want the machine to process and analyze large amount of natural language data to accomplish very interesting tasks (such as Sentiment Analysis, Language Identification and many others). Nowadays, these tasks are plenty and many works on them has been already carried out \footnote{Here is a very interesting repository collecting several NLP tasks along with their main references: \\ \url{https://github.com/Kyubyong/nlp_tasks}}. In this paper, we focus on the task of predicting whether two questions from an online forum are semantically equivalent, i.e. they can be answered by the exact same answer. This can be a very useful tool for Question-Answering community sites as they typically want to keep their databases as less redundant as possible, so to make the searches of their users both faster and more effective. Furthermore, they want to keep people from wasting their time in answering already solved questions. \\
The main contribution of this work is \cite{bogdanova2015detecting}. In their work, Bogdanova et al. propose a simple Convolutional Neural Network (CNN) to detect semantically equivalent questions. The proposed network first transforms words into \emph{word embeddings} and then applies a convolutional layer which generates question-wide vectors representations. Finally, pairs of questions are scored using a similarity function. The most interesting experimental result is the evaluation of in-domain word embeddings versus the word embeddings generated using all of the English Wikipedia. In particular, they shown that domain-specific word embeddings achieve higher performances. \\

In general, the results obtained by our implementation are very close to the ones of the replicated paper. Surprisingly, as we will discuss in Section 4, the main differences in the results are about the optimal hyper-parameter values of the network.\\
The authors of the paper accurately described the architecture of the network, but many crucial details of the model are missing in their description. Hence, we have found quite stimulating to replicate this paper filling the holes with our own ideas. In particular, our efforts focused mainly on the following three aspects that were not explicitly treated in the paper: 
    \paragraph{Handling questions of different sizes.}
    Bogdanova et al. explicitly state that the different sizes of questions is one of the main challenges of their work. Nevertheless, they do not specify the way they solved this issue. We, therefore, tried to fix the length of each question to the \textbf{maximum} and to the \textbf{mean} length of all the questions. The first approach resulted to yield the greater accuracy.
    \paragraph{The type of convolution.}
    For the implementation of our network we used a 1-dimensional convolution. Indeed, text can be seen as images with the input maps having only one dimension (height, without width) corresponding to the question length and where the number of input maps is nothing but the size of the word embeddings. The type of convolution represents another missing detail of \cite{bogdanova2015detecting}.    
    \paragraph{The training of in-domain word embeddings.}
    The paper does not provide any information about the model used to train in-domain word embeddings. Initially, we tried to build our own \emph{word2vec} model \cite{mikolov2013distributed} to generate word-embeddings using as training data the same dataset used to train the main network \footnote{To do so we followed this good tutorial: \url{http://adventuresinmachinelearning.com/word2vec-keras-tutorial/}}. However, the training of this network turned out to be very slow and resulting embeddings perfomed poorly. \\
    We, then, decided to use Gensim, a very simple library which generates word-embeddings given the training data. The resulting embeddings performed very well, allowing the main network to achieve up to 90\% of accuracy on the validation set.

@@

%-------------------------------------------------------------------------
\subsection{Question of different length}
As starting point let's understand why unequal text length is a problem in NLP. The basic recipe for processing text in neural models is as follows:
\begin{itemize}\itemsep0.1pt
\item Convert text to tokens
\item Convert tokens to token IDs
\item Look up embedding vectors for IDs and construct a 2D tensor that are input to subsequent layers.
\end{itemize}•
Now, most neural network APIs (such as Keras) are “define-and-run” so we cannot change the dimensions of various matrices and tensors during runtime. All the techniques that can be used to tackle this issue force us to introduce new hyper-parameters or work-around in our model which are not fundamental for  the NLP task but are present to only handle unequal text lengths.\\
We decided to try two of these techniques, which we found out to be the best practices:
\begin{itemize}
	\item Fixing the length of all the questions to the \textbf{maximum} length. All the questions that are shorter are padded using the \emph{UNKNOWN} word (index $0$ of the vocabulary).
	\item Fixing the length of all the questions to the \textbf{mean} length. All the questions that are shorter are padded as in the case above; all the questions that are longer are chopped.
\end{itemize}
To keep the architecture of our network as simpler as possible we decided to apply this techniques as a preprocessing phase, without adding any further layer to the network architecture. Therefore, our network accepts as input only questions of the same length.\\
The experimental results show that the \emph{max} approach provides the best performances. This is probably due to the fact that the \emph{mean} approach is very likely to drop important semantic information from the sentence. It is worth mentioning that, in order for the \emph{max} approach to perform at its best, it can be a good choice dropping the outliers, i.e. the questions that are too long with respect to the other observations.
%-------------------------------------------------------------------------
\subsection{Type of convolution}

To implement the convolution described in Section 3.1.2., we decided to use a 1-dimensional convolutional layer. The reason behind this choice is that, as images are seen as three input maps (R,G,B) where each map is a 2-dimensional array (matrix) with height and width, similarly text can be seen as $d$ input maps where each map has only one dimension which is the length of the question. This interpretation of the 1-dimensional convolution is exemplified in Figure \ref{fig:conv}.\\
After applying vertically the sum operation to the output of this layer, the resulting vector is exactly the one described by Equation \ref{eq:conv}

@@

\paragraph{Datasets.}
Our dataset is from Ask Ubuntu Community Questions and Answers site \footnote{http://askubuntu.com}. Ask Ubuntu is a community for Ubuntu users and developers, and it is part of the Stack Exchange Q\&A communities. The users of these communities can ask and answer ques- tions, and vote up and down both questions and answers. Users with high reputation become moderators and can label a new question as a duplicate to an existing question. Usually it takes five votes from different moderators to close a question or to mark it as a duplicate. 
We use Ask Ubuntu data dump \footnote{ \url{https://archive.org/download/stackexchange}} provided in April 2018, that contains posts from 2010 to March 2018. From the data dump we have selected only the posts that are questions, the amount of questions is 285651. We extract all question pairs linked as duplicated and from those we consider only 15500 pairs in order to be able to compare the obtained results with the ones from the paper \cite{bogdanova2015detecting}. Afterward we have randomly generated the same number of non duplicate pair of posts. \\
For our experiments, we divide the dataset in the same way of the paper \cite{bogdanova2015detecting}: 24K pairs for the training set, 6K pairs for the test set and 1K pairs for the validation set.

\paragraph{Tokenization.}
Tokenization is the process of breaking up a given text, in this case every question, into units called tokens. The tokens may be words or number or punctuation mark: we decide to consider only words and to disregard those tokens that are punctuation, symbols, tabulation, white spaces, articles and prepositions.
For each word we took only its lemma. Lemmatization is a morphological analysis of words, aiming to remove inflectional forms and to return the base or dictionary form of a word, which is known as the lemma.

Then we noticed that some patterns are very frequent but not so relevant in their different form. Therefore we had some other precautions:
\begin{itemize}
	\item html tags that are \texttt{code} tags are replaced by the string \texttt{thistokeniscode}
	\item html tags that are \texttt{a} tags are replaced by the string \texttt{thistokenislink}
	\item tokens that match with a time format are replaced by the string \texttt{thistokenistime}
	\item tokens that match with a date format are replaced by the string \texttt{thistokenisdate}
	\item tokens that match with a version format are replaced by the string \texttt{thistokenisversion}
	\item tokens that match with a path format are replaced by the string \texttt{thistokenispath}
	\item tokens that match with a hexadecimal format are replaced by the string \texttt{thistokenishexadecimal}
 \end{itemize}
